---
title: "PCAPlantSynergy"
author: "Erola Fenollosa"
date: "2024-02-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Build a PCA

## 1. Import the dataset and libraries
It must be stored at the same folder of the R code.
```{r}
# Remember to first do install.packages()
library(tidyverse) # Data Handling
library(GGally) # Correlation plots
library(factoextra) # PCA visualizations

pizza <- read.csv("Pizza.csv")
head(pizza)
```

The dataset was downloaded from this link: https://data.world/sdhilip/pizza-datasets and it contains information on the micronutrients of various pizzas.

The variables in the data set are:

brand - Pizza brand (class label)
id - Sample analised
mois - Amount of water per 100 grams in the sample
prot - Amount of protein per 100 grams in the sample
fat - Amount of fat per 100 grams in the sample
ash - Amount of ash per 100 grams in the sample
sodium - Amount of sodium per 100 grams in the sample
carb - Amount of carbohydrates per 100 grams in the sample
cal - Amount of calories per 100 grams in the sample

## 2. Exploratory Data Analysis
```{r}
summary(pizza) 
```

```{r}
table(pizza$brand) 
```

Missing values in the data set:
```{r}
apply(pizza[1:length(colnames(pizza))],2, anyNA)
```
Delete the variable called "id"
```{r}
pizza2 <- subset(pizza, select = -c(id))
head(pizza2)
```
```{r}
pizza2 %>% 
  gather("micronutritions", "value", 2:8) %>% 
  ggplot(aes(x = brand, y = value, fill = brand)) +
    geom_boxplot() +
    ggtitle("Histograms of micronutritions") +
    xlab("Brands") +
    ylab("Values") +
    theme_bw() +
    theme(plot.title = element_text(hjust = 0.5, size = 14)) +
    facet_wrap(~ micronutritions, scales = "free")

cor_pizza <- cor(pizza[, 2:length(pizza)])
ggcorr(cor_pizza, label = T, label_round = 2)
```

## 3. PCA

### 3.1. Variance explained
```{r}
basic_PCA <- prcomp(pizza[,-1], center = TRUE, scale = TRUE)
summary(basic_PCA)
```
```{r}
fviz_screeplot(basic_PCA, addlabels = TRUE)
```

### 3.2. Variables contribution to each component
```{r}
fviz_pca_var(basic_PCA, col.var = "contrib", repel = TRUE, axes = c(1, 3)) +
  labs(title="Variables loading for PC1 and PC2", x="PC1", y="PC2")
```

```{r}
fviz_contrib(basic_PCA, choice = "var", axes = 1)
fviz_contrib(basic_PCA, choice = "var", axes = 2)
```

```{r}
# Results for Variables
res.var <- get_pca_var(basic_PCA)
round(res.var$contrib,4)        # contains the contributions (in percentage) of the variables to the principal components.
```




### 3.3 PCA visualization
```{r}
fviz_pca_ind(basic_PCA)

fviz_pca_biplot(basic_PCA, 
                col.var = "#2E9FDF", # Variables color
                col.ind = "#696969")  # Individuals color               


fviz_pca_ind(basic_PCA,
             geom.ind = "point", # show points only (nbut not "text")
             col.ind = pizza2$brand, # color by groups
             palette = c("#00AFBB", "#E7B800", "#FC4E07","#2E9FDF", "#696969", "black", "steelblue", "pink", "coral", "lightgreen"),
             addEllipses = TRUE, # Concentration ellipses
             legend.title = "Groups") + scale_shape_manual(values=c(19, 19, 19, 19, 19, 19, 19, 19, 19, 19))

fviz_pca_ind(basic_PCA,
             geom = "point", # show points only (nbut not "text")
             col.ind = pizza2$brand, # color by groups
             palette = c("#00AFBB", "#E7B800", "#FC4E07","#2E9FDF", "#696969", "black", "steelblue", "pink", "coral", "lightgreen"),
             addEllipses = TRUE, ellipse.type = "confidence",
             legend.title = "Groups") + scale_shape_manual(values=c(19, 19, 19, 19, 19, 19, 19, 19, 19, 19))


```

